---
title: "SB_project"
author: "MariaRomanova"
output: pdf_document
---
# All required packages
```{r}
ipak <- function(pkg){
  new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
  if (length(new.pkg)) 
    install.packages(new.pkg, dependencies = TRUE)
  sapply(pkg, require, character.only = TRUE)
}

required_libraries = c('sas7bdat','tidyr','plyr','reshape2','MASS','ggplot2','UpSetR','data.table','Hmisc','foreach','dplyr')
ipak(required_libraries)
```

# All datasets needed
```{r}
analysis <- data.frame(sas7bdat::read.sas7bdat('Baseline/assay.sas7bdat'))
oldtnfa <- data.frame(sas7bdat::read.sas7bdat('Baseline/oldtnfa.sas7bdat'))
analysis <- join(analysis,oldtnfa)
analysis_2 <- data.frame(sas7bdat::read.sas7bdat('Follow_up1/assay.sas7bdat'))
analysis_3 <- data.frame(sas7bdat::read.sas7bdat('Follow_up2/assay.sas7bdat'))
analysis_4 <- data.frame(sas7bdat::read.sas7bdat('Follow_up3/assay.sas7bdat'))
analysis_5 <- data.frame(sas7bdat::read.sas7bdat('Follow_up4/assay.sas7bdat'))
diseases <- data.frame(sas7bdat::read.sas7bdat('Baseline/deseases.sas7bdat'))
interview <- data.frame(sas7bdat::read.sas7bdat('Baseline/interview.sas7bdat'))
physical <- data.frame(sas7bdat::read.sas7bdat('Baseline/phisical.sas7bdat'))
under_65 <- subset(analysis,analysis$X_AGEL<65)
under_65_M <- subset(under_65,under_65$SEX==1) %>% dplyr::select(-SEX)
under_65_F <- subset(under_65,under_65$SEX==2) %>% dplyr::select(-SEX)
```
# # PART 1. Separate study
# Oxidized guanine, guanosine and deoxyguanosine using Baseline 24-hour urine sample
X_CASCTL 1=case,0=control X_UGUO24  X_UXGA24 
+ lm
```{r}
separate_study <- c('X_CASCTL','X_PPAIR','X_UGUA24','X_UXGA24','X_UGUO24','X_UXGO24','X_UDG24','X_UXDG24','X_UCOR24')
separate_study_df_M <- subset(analysis,analysis$SEX==1) %>%
  dplyr::select(separate_study)
separate_study_df_F <- subset(analysis,analysis$SEX==2) %>% 
  dplyr::select(separate_study)

lst <- list(separate_study_df_M,separate_study_df_F)

lapply(lst, function(x){
  df <- x
  case    <- subset(df, df$X_CASCTL == '1') %>% 
    dplyr::select(-c('X_CASCTL','X_PPAIR'))
  control <- subset(df, df$X_CASCTL == '0') %>% 
    dplyr::select(-c('X_CASCTL','X_PPAIR'))
  
    corrplot::corrplot(cor(case,method = 'spearman',
                         use='pairwise.complete'),
                     method = c("circle"),
                     type='upper')
    corrplot::corrplot(cor(control,method = 'spearman',
                         use='pairwise.complete'),
                     method = c("circle"), 
                     type='upper')

  
  l <- names(case)
  lapply(l, function(x){
  return(list(x,summary(lm(X_UCOR24~case[,x],data=case))))})
    summary(lm(X_UCOR24~X_UGUA24+X_UXGO24,data=case))
    summary(lm(X_UCOR24~X_UGUA24+X_UGUO24,data=case))
})
```

```{r}
case    <- subset(separate_study_df_M, separate_study_df_M$X_CASCTL == '1') %>% 
    dplyr::select(-c('X_CASCTL','X_PPAIR'))
df <- case
sample <- sample.int(n=nrow(df),size=floor(.75*nrow(df)))
train <- df[sample,]
test <- df[-sample,]
new_mod <- lm(X_UCOR24~X_UXGO24,data=train)
summary(new_mod)
pred <- predict(new_mod,test)
ggplot(train,aes(x=X_UXGO24,y=X_UCOR24))+
  geom_point() +
  geom_smooth(method = 'lm') +
  geom_point(data=test,col='green') +
  geom_point(data = new_mod,aes(y=new_mod$fitted.values),color='red') 
```

# Part 2
Finding all significant correlations with spearman, creating a list of correlations dataframes over all 5 waves of study
```{r}
qplot(analysis$X_AGEL,
      geom="histogram",
      binwidth = 1,  
      main = "Histogram of Age", 
      xlab = "Age",
      ylab = "Count",
      col=I("red"),
      fill=I("lightblue")) + 
  scale_x_continuous(breaks=c(20,40,65,80,100)) 
```

Only people up to 65 were taken
```{r}
# List with 5 data frames of significant correlations
lst <- list(analysis,analysis_2,analysis_3,analysis_4,analysis_5)
cor <- lapply(lst, function(x){
  
  under_65 <- subset(x,x[,6]<65) %>% select_if(is.numeric)
  
  under_65_M <- subset(under_65,under_65$SEX==1) %>% dplyr::select(-SEX)
  
  under_65_F <- subset(under_65,under_65$SEX==2) %>% dplyr::select(-SEX)
  
  flat_cor_mat <- function(cor_r, cor_p){
    library(tidyr)
    library(tibble)
    cor_r <- rownames_to_column(as.data.frame(cor_r), var = "row")
    cor_r <- gather(cor_r, column, cor, -1)
    cor_p <- rownames_to_column(as.data.frame(cor_p), var = "row")
    cor_p <- gather(cor_p, column, p, -1)
    cor_p_matrix <- left_join(cor_r, cor_p, by = c("row", "column"))
    cor_p_matrix %>%
      subset(cor > 0.5 | cor <(-0.5))
  }
  
  under_65_M <- Hmisc::rcorr(as.matrix(under_65_M),type='spearman')
  under_65_F <- Hmisc::rcorr(as.matrix(under_65_F),type='spearman')
  
  cor_matrix_1 <- flat_cor_mat(under_65_M$r, under_65_M$P) %>%
    subset(cor> 0.5 | cor<(-0.5) & p < 0.05) %>% subset(!duplicated(cor))
  
  cor_matrix_2 <- flat_cor_mat(under_65_F$r, under_65_F$P) %>%
    subset(cor> 0.5 | cor<(-0.5) & p < 0.05) %>% subset(!duplicated(cor))
  
  cor_matrix_1$cor_pair <- paste(cor_matrix_1$row,cor_matrix_1$column,sep=',')
  cor_matrix_2$cor_pair <- paste(cor_matrix_2$row,cor_matrix_2$column,sep=',')
  
  return(merge(cor_matrix_1,cor_matrix_2, by='cor_pair',all=T))
})
```
Interesting correlation of total cholesterol and vitamin E
```{r}
library(ggpubr)
ggplot(under_65_F,aes(under_65_F$X_COLTOT,under_65_F$X_ATOCRS)) +
  geom_point() +
  geom_smooth(method = 'lm') +
  stat_cor(method = "spearman",label.x = 50) +
  theme_light()
```

```{r}
df <- under_65_F
sample <- sample.int(n=nrow(df),size=floor(.75*nrow(df)))
train <- df[sample,]
test <- df[-sample,]
new_mod <- lm(X_COLTOT~X_ATOCRS,data=train)
summary(new_mod)
pred <- predict(new_mod,test)
ggplot(train,aes(x=X_ATOCRS,y=X_COLTOT))+
  geom_point() +
  geom_smooth(method = 'lm') +
  geom_point(data=test,col='green') +
  geom_point(data = new_mod,aes(y=new_mod$fitted.values),color='red') +
  annotate("text", x=30, y=400, label= "Adjusted R-squared:  0.5173") +
  annotate("text", x=30, y=380, label= "lm(formula = X_COLTOT ~ X_ATOCRS") +
  ggtitle('Linear model:test/train. Total cholesterol and vitamin E')
```

# Age prediction with linear model
Males
```{r}
library(randtests)
df <- select(under_65_M,c(X_AGEL,X_TIGF1,X_COLTOT,X_DHEAS,X_BIOATS,X_ATOCRS,X_FREETS,X_TSSHBG))
corrplot::corrplot(cor(df,use='pairwise.complete'),p.mat=corrplot::cor.mtest(df)$p)
l <- names(df)
lapply(l, function(x){
  return(list(x,summary(lm(X_AGEL~df[,x],data=df))))})
fit <- lm(X_AGEL~X_TIGF1+X_DHEAS+X_BIOATS,data=df)
summary(fit)
runs.test(fit$residuals)

sample <- sample.int(n=nrow(df),size=floor(.75*nrow(df)))
train <- df[sample,]
test <- df[-sample,]
new_mod <- lm(X_AGEL~.,data=train)
summary(new_mod)
plot(new_mod)
library(randtests)
runs.test(new_mod$residuals)
```
Females
```{r}
df <- select(under_65_F,c(X_AGEL,X_TIGF1,X_COLTOT,X_DHEAS,X_ESTDIO,X_ATOCRS,X_FREETS,X_TSSHBG))
corrplot::corrplot(cor(df,use='pairwise.complete'),p.mat=corrplot::cor.mtest(df)$p)
l <- names(df)
lapply(l, function(x){
  return(list(x,summary(lm(X_AGEL~df[,x],data=df))))})
fit <- lm(X_AGEL~X_TIGF1+X_DHEAS,data=df)
summary(fit)
plot(fit)
runs.test(fit$residuals)

sample <- sample.int(n=nrow(df),size=floor(.75*nrow(df)))
train <- df[sample,]
test <- df[-sample,]
new_mod <- lm(X_AGEL~.,data=train)
summary(new_mod)
plot(new_mod)
library(randtests)
runs.test(new_mod$residuals)
```

# Part 3 Association rules
```{r}
library(arules)
physical <- select(physical,c(PX1_V20,PX1_V21,CODE98))
smoke <- select(interview,c(CODE98,IX9_V37))
vec1 <- intersect(physical$CODE98,analysis$CODE98) %>%
  intersect(diseases$CODE98)
analysis <- analysis[analysis$CODE98 %in% vec1,]
df <- select(analysis,c(X_U_PH,X_U_PRO, X_COLTOT,X_A_G,X_AGEL,X_GLU,X_TIGF1,X_INSULN,X_IL6,X_A_G,X_TNFA,XBIOATSM,X_ESTDIO,X_TNFAR1,X_LUTEIN,X_FIBRIN,X_TRIGLI,X_UBPA24,X_PALK,X_ATOCRS,X_OM6_A,X_DHEAS,X_MIO,X_COLHDL,X_COLLDL,X_UCRE24,X_UGUA24,X_UCOR24))
df <- as.data.frame(sapply(names(df), function(x){
  discretize(df[,x],method = 'cluster')
  }))
vec <- c('AXCANCER','AXEPATOC','AXMI','AXCHF','AXSTROKE','AXPARK','AXBPCO','AXANGINA','AXPADI','AXASTHMA','AXMIPERT','AXMWAIST','AXMHTRIG','AXMHGLUC','AXMLOHDL','AXDEMENT','AXVASDEM','AXMETBOL','SEX','CITE')
 
diseases <- diseases[diseases$CODE98 %in% vec1,]
physical <- mutate(physical,BMI=PX1_V20/((PX1_V21/100)^2))

smoke <- smoke[smoke$CODE98 %in% vec1,]
df$BMI <- physical$BMI
diseases_ <- diseases[diseases$CODE98 %in% vec1,names(diseases) %in% vec]
df$smoke <- smoke$IX9_V37
df <- cbind(df,diseases_)
df <- df%>%
    mutate_if(is.numeric,as.factor)
tData <- as (df, "transactions")
rules <- apriori(tData, parameter = list(supp = 0.01, conf = 0.8, target = "rules", maxlen=2))
inspect(head(rules, by = "lift",n=1000))
```



















